---
title: "Assessment R Markdown"
author: "B203349"
date: "16/06/2022"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


[Click here to link to my git hub repository](https://github.com/B203349/B203349_assessment)

# Loading the required packages required for this script
```{r a,message = FALSE, warning = FALSE}
library(dataMeta)
library(tidyverse)
library(here)
library(NHSRdatasets)
library(knitr)
library(scales)
library(lubridate)
library(caret)
```

# Loading NHS Datasets 

## Data
The **NHS England accident and emergency attendances and admissions (ae_attendances)** dataset will be loaded from the NHSRdatasets package. This dataset reports on the attendances, four-hour breaches and admissions for all A&E departments in England for the years 2016/17 through 2018/19 (Apr-Mar)

## Loading ae_attendances data
ae_attendances data is loaded from the NHSdatasets package
```{r g,message = FALSE, warning = FALSE}
data(ae_attendances)
ae<-ae_attendances
```

## Viewing the data
The `head()` function is used to look at top n rows of a data frame (default n = 6 rows)
```{r k,message = FALSE, warning = FALSE}
head(ae)
```
### Check for missing data
Calculate how many NAs there are in each variable
```{r n,message = FALSE, warning = FALSE}
ae %>% 
  map(is.na) %>%
map(sum)
```
Data is complete

## Add an index link column to ae_attendances data  
The `rowid_to_column()` function is used to convert row identities to a column named index. This facilitates the linking of partitioned datasets to the raw data if required in the future.
```{r o,message = FALSE, warning = FALSE}
ae <- rowid_to_column(ae, "index")
```


### Preview dataset
Tabulate data, and create report with `kable()`
```{r p,message = FALSE, warning = FALSE}
ae %>%
  mutate_at(vars(period), format, "%b-%y") %>% 
  mutate_at(vars(attendances, breaches, admissions), comma) %>%
  head(10) %>%
  kable()
```



## Save the raw ae_attendances data to your 'RawData' folder
```{r q,message = FALSE, warning = FALSE}
write_csv(ae, here("RawData", "ae_attendances.csv"))
```

## Subsetting the data
The aim of my data capture tool is to help identify performance outliers with regard to the 4hr breaches. In order to do this effectively I will need use the entire dataset (i.e. the variables: index, period, attendances, breaches, organisation name, and organisation type).

Of note the **performance** variable was deliberately not created as this is a function of the **attendances** and **breaches** variables. The variable can be created after the collection of data at the data analysis stage. 


## Creating test and training datasets

### Calculate the proportion (`prop`) of the raw data to assign to the training data.
The proportion of the raw that needs to be assigned to the training data to ensure there is only 10 to 15 records in the test data is:
```{r y,message = FALSE, warning = FALSE}
prop<-(1-(15/nrow(ae)))
print(prop)
```

### Splitting the raw data
The createDataPartition() function from the *caret* package is used to splint the raw data into test and training data sets.

The `set.seed()` function allows a random numner to be generated in a reproducable fashion. This ensures that every time the script is run the raw data will be partitioned into the same test and training datasets.
```{r z,message = FALSE, warning = FALSE}
set.seed(333)
trainIndex <- createDataPartition(ae$index, p = prop, 
                                  list = FALSE, 
                                  times = 1)
head(trainIndex)
```

Assign all records in the trainIndex to the training data
```{r zz,message = FALSE, warning = FALSE}
aeTrain <- ae[ trainIndex,]
nrow(aeTrain)
```

Save the training data to the to working data folder 'Data' as ae_attendances_train.csv
```{r ab,message = FALSE, warning = FALSE}
write_csv(aeTrain, here("Data", "ae_attendances_train.csv"))
```

Extract the test data (all records that are not in the trainIndex (`-trainIndex`) are assigned to the test data).
```{r ac,message = FALSE, warning = FALSE}
aeTest  <- ae[-trainIndex,]
nrow(aeTest)
```

Set aside the fist record for markers to test and evaluate the data-capture tool.
```{r ad,message = FALSE, warning = FALSE}
aeTestMarker  <- aeTest[1,]
```

Save the marker test data to the to working data folder 'Data' as ae_attendances_test_marker.csv
```{r af,message = FALSE, warning = FALSE}
write_csv(aeTestMarker, here("Data", "ae_attendances_test_marker.csv"))
```


The remaining records are set aside to test with the data capture tool. These are saved to the to working data folder 'Data' as ae_attendances_test.csv
```{r ag,message = FALSE, warning = FALSE}
aeTest  <- aeTest[2:nrow(aeTest),]
write_csv(aeTest, here("Data", "ae_attendances_test.csv"))
```

# Data Dictionary for test data

## Reading the collected data
`read_csv()` function from the *readr* package is used to read my collected data from the Raw data folder.
```{r data}
CollectedData=read_csv(here("RawData", "CollectedDataFinal.csv"))
```

## Building a linker data frame

### Variable descriptions
String vectors are created for the different variable descriptions

```{r variable_description}
variable_description <- c("The index column that allows us to link the data collected to the original ae_attendances data in the 'RawData' folder.",
"The month that this activity relates to, stored as a date (1st of each month).",
"The Organisation data service (ODS) code for the organisation. If you want to know the organisation associated with a particular ODS code, you can look it up from the following address: https://odsportal.digital.nhs.uk/Organisation/Search.", 
"The department type for this activity.",
"The number of attendances for this department type at this organisation for this month.", 
"The number of attendances that breached the four-hour target.",  
"The number of attendances that resulted in an admission to the hospital.",
"The consent from the end-user to process and share the data collected with the data capture tool.")
print(variable_description)
```

### Variable types
A string vector is created representing the different variable types. It is a vector of integers with values 0 or 1. 0 is used for a variable with quantitative values (measured values) variables and 1 for fixed values (allowable values or codes) variables.

`glimpse()` function from *tibble* package is used to view the variable types in the CollectedData data frame
```{r glimpse2}
glimpse(CollectedData) 
```

This indicated that there are four quantitative values (measured values) variables and four fixed values (allowable values or codes) variables.
```{r variable_type}
variable_type <- c(0, 1, 1, 1, 0, 0, 0, 1)
print(variable_type)
```

### Building the linker
`build_linker()` function from the *dataMeta* package is used to construct an intermediary (linker) data frame between the CollectedData and the data dictionary. 
```{r build_linker}
linker<-build_linker(CollectedData, variable_description, variable_type)
print(linker)
```

## Data dictionary
the `build_dict()` function from the *dataMeta* package is used to construct a data dictionary for a CollectedData data frame with the aid of the linker data frame.
```{r dictionary}
dictionary <- build_dict(my.data = CollectedData, linker = linker)
```
```{r notes}
dictionary[6,4]<-"C82010: Prescribing Cost Centre - OAKHAM MEDICAL PRACTICE"
dictionary[7,4]<-"RDZ: NHS Trust - THE ROYAL BOURNEMOUTH AND CHRISTCHURCH HOSPITALS NHS FOUNDATION TRUST"
dictionary[8,4]<-"RVR: NHS Trust - EPSOM AND ST HELIER UNIVERSITY HOSPITALS NHS TRUST"
dictionary[9,4]<-"RQM: NHS Trust - CHELSEA AND WESTMINSTER HOSPITAL NHS FOUNDATION TRUST"
dictionary[10,4]<-"R1F: NHS Trust - ISLE OF WIGHT NHS TRUST"
dictionary[11,4]<-"RE9: NHS Trust - SOUTH TYNESIDE NHS FOUNDATION TRUST"
dictionary[12,4]<-"RNL: NHS Trust - NORTH CUMBRIA UNIVERSITY HOSPITALS NHS TRUST"
dictionary[13,4]<-"RJ1: NHS Trust - GUY'S AND ST THOMAS' NHS FOUNDATION TRUST"
dictionary[14,4]<-"RKB: NHS Trust - UNIVERSITY HOSPITALS COVENTRY AND WARWICKSHIRE NHS TRUST"
dictionary[15,4]<-"NLO12: Independent Sector H/c Provider Site - OAKHAM URGENT CARE CENTRE"
dictionary[26,4] <-"other: Other types of A&E/minor injury activity with designated accommodation for the reception of accident and emergency patients."
dictionary[27,4]<- "1: Emergency departments are a consultant-led 24-hour service with full resuscitation facilities and designated accommodation for the reception of accident and emergency patients."
dictionary[28,4] <- "2: Consultant-led mono speciality accident and emergency service (e.g. ophthalmology, dental) with designated accommodation for the reception of patients. other: Other types of A&E/minor injury activity with designated accommodation for the reception of accident and emergency patients."
```

### Save the data dictionary for CollectedData to the 'RawData' folder
```{r w,message = FALSE, warning = FALSE}
write_csv(dictionary, here("RawData", "CollectedData_DataDictionary.csv"))
```
## Appending the data dictionary to data
Incorporate attributes as metadata to the CollectedData as metadata using the `incorporate_attr()` function from the *dataMeta* package.
This requires the CollectedData, dictionary, and main_string as inputs (main_string is a character string describing the CollectedData data frame)

### Create main_string for attributes
```{r main_string}
main_string <- "This data describes the NHS England accident and emergency (A&E) attendances and breaches of four-hour wait time target data from the *NHSRdatasets* package collected by the data capture tool."
```
### Incorporate attributes as metadata
```{r complete_dataset}
complete_CollectedData <- incorporate_attr(my.data = CollectedData, data.dictionary = dictionary,
main_string = main_string)
attributes(complete_CollectedData)$author[1]<-"B203349"
complete_CollectedData
attributes(complete_CollectedData)
```

### Save the CollectedData with attributes
```{r save_it}
save_it(complete_CollectedData, here("RawData", "complete_CollectedData"))
```

# Data Capture Tool (Python)

A data capture tool was created. The code was written in Python using a Jupyter notebook. Data will be collected using an interactive graphic user interface that was built using widgets from `ipywidgets` package. Collected data will be stored 

